---
title: "Contact helps dispreferred combinations of typological features to survive: geospatial evidence"
subtitle: "Statistical analyses"
output: html_document
params:
  dataset: null
---


# Preliminaries

```{r}
# required packages
require(tidyverse)
require(reshape2)
require(broom)
require(pixiedust)
require(emmeans)
source("pvalString_demo.R")

# create results/tables, if it doesn't exist already
try(dir.create("../../results/tables", recursive=TRUE))

# load data
dataset <- params$dataset
#wals <- read.csv("../results/wals/results_combined.csv")
wals <- read.csv(paste0("../../results/", dataset, "/results_combined.csv"))

# make "control" the reference level of "status" factor
wals$status <- factor(wals$status)
wals$status <- relevel(wals$status, ref="non-interacting")

# absolute value of corrected correlation coefficient
wals$abs_corrected_phi <- abs(wals$corrected_phi)

# restrict to one single neighbourhood size
wals$k <- wals$degree
wals <- wals[wals$k == 10, ]
```

# Descriptive statistics

Mean sample size:

```{r}
round(mean(wals$N))
```
This suggests the following neighbourhood size to use:

```{r}
round(sqrt(round(mean(wals$N))))
```

# Basic model: comparison of $\Delta$ between typologies of different statuses

## Under-represented types ($\Delta^-$)

```{r}
mod1_under_w <- lm(Delta_under ~ status + skewness, data=wals)

mathematize <- function(value) {
  paste0("[", value, "]")
}

mod1_under_w %>% 
  dust %>% 
  sprinkle(round=3) %>% 
  sprinkle(cols="term", replace=c("intercept", "interacting (vs. control)", 
                                  "unknown (vs. control)", "TAS")) %>% 
  sprinkle(cols="p.value", fn = quote(pvalString_demo(value, stars=TRUE))) %>%
  write.table(file="../../results/tables/mod1_cat_under_wals.csv", 
              row.names=FALSE, col.names=FALSE, sep=",", quote=FALSE)

print(summary(mod1_under_w))
```

```{r}
walsh <- wals[wals$status != "unknown", ]

walsh <- melt(walsh, id.vars=c("pair", "status", "skewness"), measure.vars=c("Delta_over", "Delta_under"))

mod <- lm(value ~ variable*status, walsh)

emm <- emmeans(mod, specs = pairwise ~ variable:status)
con <- as.data.frame(emm$contrasts)
eff <- as.data.frame(eff_size(emm, sigma=sigma(mod), edf=mod$df, method="identity"))

print(con)
print(eff)

print(-con[con$contrast == "(Delta_under non-interacting) - Delta_under interacting", ]$estimate)
print(con[con$contrast == "(Delta_under non-interacting) - Delta_under interacting", ]$p.value)
print(-eff[eff$contrast == "((Delta_under non-interacting) - Delta_under interacting)", ]$effect.size)

print(-con[con$contrast == "(Delta_over non-interacting) - Delta_over interacting", ]$estimate)
print(con[con$contrast == "(Delta_over non-interacting) - Delta_over interacting", ]$p.value)
print(-eff[eff$contrast == "((Delta_over non-interacting) - Delta_over interacting)", ]$effect.size)
```


## Over-represented types ($\Delta^+$)

```{r}
mod1_over_w <- lm(Delta_over ~ status + skewness, data=wals)

mod1_over_w %>% 
  dust %>% 
  sprinkle(round=3) %>% 
  sprinkle(cols="term", replace=c("intercept", "interacting (vs. control)", 
                                  "unknown (vs. control)", "TAS")) %>% 
  sprinkle(cols="p.value", fn = quote(pvalString_demo(value, stars=TRUE))) %>%
  write.table(file="../../results/tables/mod1_cat_over_wals.csv", 
              row.names=FALSE, col.names=FALSE, sep=",", quote=FALSE)

print(summary(mod1_over_w))
```


# Model 2: continuous predictor

## Under-represented types ($\Delta^-$)

```{r}
mod1_under_w <- lm(Delta_under ~ abs_corrected_phi + skewness, data=wals)

mathematize <- function(value) {
  paste0("[", value, "]")
}

mod1_under_w %>% 
  dust %>% 
  sprinkle(round=3) %>% 
  sprinkle(cols="term", replace=c("intercept", "|φ_c|", "TAS")) %>% 
  sprinkle(cols="p.value", fn = quote(pvalString_demo(value, stars=TRUE))) %>%
  write.table(file="../../results/tables/mod1_cts_under_wals.csv", 
              row.names=FALSE, col.names=FALSE, sep=",", quote=FALSE)

print(summary(mod1_under_w))
```

## Under-represented types ($\Delta^-$), only interacting typologies

```{r}
mod1_under_w <- lm(Delta_under ~ abs_corrected_phi + skewness, data=wals[wals$status == "interacting", ])

mathematize <- function(value) {
  paste0("[", value, "]")
}

mod1_under_w %>% 
  dust %>% 
  sprinkle(round=3) %>% 
  sprinkle(cols="term", replace=c("intercept", "|φ_c|", "TAS")) %>% 
  sprinkle(cols="p.value", fn = quote(pvalString_demo(value, stars=TRUE))) %>%
  write.table(file="../../results/tables/mod1_cts_inter_under_wals.csv", 
              row.names=FALSE, col.names=FALSE, sep=",", quote=FALSE)

print(summary(mod1_under_w))
```



## Over-represented types ($\Delta^+$)

```{r}
mod1_over_w <- lm(Delta_over ~ abs_corrected_phi + skewness, data=wals)

mod1_over_w %>% 
  dust %>% 
  sprinkle(round=3) %>% 
  sprinkle(cols="term", replace=c("intercept", "|φ_c|", "TAS")) %>% 
  sprinkle(cols="p.value", fn = quote(pvalString_demo(value, stars=TRUE))) %>%
  write.table(file="../../results/tables/mod1_cts_over_wals.csv", 
              row.names=FALSE, col.names=FALSE, sep=",", quote=FALSE)

print(summary(mod1_over_w))
```



## Over-represented types ($\Delta^+$), interacting typologies only

```{r}
mod1_over_w <- lm(Delta_over ~ abs_corrected_phi + skewness, data=wals[wals$status == "interacting", ])

mod1_over_w %>% 
  dust %>% 
  sprinkle(round=3) %>% 
  sprinkle(cols="term", replace=c("intercept", "|φ_c|", "TAS")) %>% 
  sprinkle(cols="p.value", fn = quote(pvalString_demo(value, stars=TRUE))) %>%
  write.table(file="../../results/tables/mod1_cts_inter_over_wals.csv", 
              row.names=FALSE, col.names=FALSE, sep=",", quote=FALSE)

print(summary(mod1_over_w))
```



