---
title: "Contact helps dispreferred combinations of typological features to survive: geospatial evidence"
subtitle: "Statistical analyses"
output: html_document
---


# Preliminaries

```{r}
# required packages
require(tidyverse)
require(broom)
require(pixiedust)
require(emmeans)

# create results/tables, if it doesn't exist already
try(dir.create("../results/tables", recursive=TRUE))

# load data
source("load_data.R")
```

# Descriptive statistics

Mean sample size, WALS:

```{r}
round(mean(wals$N))
```

Mean sample size, Grambank:

```{r}
round(mean(gram$N))
```

These suggest the following neighbourhood sizes to use:

```{r}
round(sqrt(round(mean(wals$N))))
```

and

```{r}
round(sqrt(round(mean(gram$N))))
```


# Basic model: comparison of $\Delta$ between typologies of different statuses

## Under-represented types ($\Delta^-$)

```{r}
mod1_under_w <- lm(Delta_under ~ status+abs(phi), data=wals)
mod1_under_g <- lm(Delta_under ~ status+abs(phi), data=gram)

mathematize <- function(value) {
  paste0("[", value, "]")
}

mod1_under_w %>% 
  dust %>% 
  sprinkle(round=3) %>% 
  sprinkle(cols="term", replace=c("intercept", "interacting (vs. non-interacting)", 
                                  "unknown (vs. non-interacting)", "|φ|")) %>% 
  write.table(file="../results/tables/mod1_under_wals.csv", 
              row.names=FALSE, col.names=FALSE, sep=",", quote=FALSE)

mod1_under_g %>% 
  dust %>% 
  sprinkle(round=3) %>% 
  sprinkle(cols="term", replace=c("intercept", "interacting (vs. non-interacting)", 
                                  "unknown (vs. non-interacting)", "|φ|")) %>% 
  write.table(file="../results/tables/mod1_under_grambank.csv", 
              row.names=FALSE, col.names=FALSE, sep=",", quote=FALSE)

print(summary(mod1_under_w))
print(summary(mod1_under_g))
```


## Over-represented types ($\Delta^+$)

```{r}
mod1_over_w <- lm(Delta_over ~ status+abs(phi), data=wals)
mod1_over_g <- lm(Delta_over ~ status+abs(phi), data=gram)

mod1_over_w %>% 
  dust %>% 
  sprinkle(round=3) %>% 
  sprinkle(cols="term", replace=c("intercept", "interacting (vs. non-interacting)", 
                                  "unknown (vs. non-interacting)", "|φ|")) %>% 
  write.table(file="../results/tables/mod1_over_wals.csv", 
              row.names=FALSE, col.names=FALSE, sep=",", quote=FALSE)

mod1_over_g %>% 
  dust %>% 
  sprinkle(round=3) %>% 
  sprinkle(cols="term", replace=c("intercept", "interacting (vs. non-interacting)", 
                                  "unknown (vs. non-interacting)", "|φ|")) %>% 
  write.table(file="../results/tables/mod1_over_grambank.csv", 
              row.names=FALSE, col.names=FALSE, sep=",", quote=FALSE)

print(summary(mod1_over_w))
print(summary(mod1_over_g))
```


# Model 2: model comparison between $\varphi$ and $\varphi_c$ as predictors

## Under-represented types

```{r}
mod_w <- lm(Delta_under ~ abs(phi), data=wals)
mod_wc <- lm(Delta_under ~ abs(corrected_phi), data=wals)

print(summary(mod_w))
print(summary(mod_wc))

print(AIC(mod_w))
print(AIC(mod_wc))

mod_g <- lm(Delta_under ~ abs(phi), data=gram)
mod_gc <- lm(Delta_under ~ abs(corrected_phi), data=gram)

print(summary(mod_g))
print(summary(mod_gc))

print(AIC(mod_g))
print(AIC(mod_gc))

#mod_w %>% dust %>% sprinkle(round=5) %>% write.csv(file="../results/tables/mod2_under_wals.csv", row.names=FALSE)
#mod_g %>% dust %>% sprinkle(round=5) %>% write.csv(file="../results/tables/mod2_under_grambank.csv", row.names=FALSE)
#mod_wc %>% dust %>% sprinkle(round=5) %>% write.csv(file="../results/tables/mod2_under_corrected_wals.csv", row.names=FALSE)
#mod_gc %>% dust %>% sprinkle(round=5) %>% write.csv(file="../results/tables/mod2_under_corrected_grambank.csv", row.names=FALSE)
```

## Over-represented types

```{r}
mod_w <- lm(Delta_over ~ abs(phi), data=wals)
mod_wc <- lm(Delta_over ~ abs(corrected_phi), data=wals)

print(summary(mod_w))
print(summary(mod_wc))

print(AIC(mod_w))
print(AIC(mod_wc))

mod_g <- lm(Delta_over ~ abs(phi), data=gram)
mod_gc <- lm(Delta_over ~ abs(corrected_phi), data=gram)

print(summary(mod_g))
print(summary(mod_gc))

print(AIC(mod_g))
print(AIC(mod_gc))
```

